---
title: "Designing a survey around fishery-dependent data"
subtitle: "FISH 507B"
author: John Best
output:
    xaringan::moon_reader:
        lib_dir: libs
        nature:
            highlightStyle: github
            countIncrementalSlides: false
---

# We need indices of abundance

- Usually derived from survey data

- Inform stock assessments

---

# Many fisheries have fishery-dependent observations

- These are not usually used if fishery-independent observations are available

- But may be far more numerous

--

- More difficult to standardize CPUE

--

- **Preferential sampling** - fishermen tend to fish where there are more fish!

---

# Can we counter preferential sampling with careful survey design?

- How does bias decrease with additional counter-preferential survey observations?

- How do catchability estimates and catchability variance estimates vary with different levels of sampling overlap?
  
$$\renewcommand{\vec}[1]{\boldsymbol{#1}}$$

```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      out.width = "100%",
                      out.height = "100%",
                      fig.align = "center")
```

---

# Operating model

```{r}
knitr::include_graphics("../figs/abundcatch_tots.svg")
```

---

## Operating model sequence

0. **Intialize** population at carrying capacity, spatial equilibrium.
1. **Harvest biomass** via survey (fixed survey stations in order) and commercial (preferential targeting) vessels.
2. **Produce** biomass via regional Beverton-Holt model.
3. **Move** toward more-preferred habitats.
4. **Repeat** from 2.

---

## Habitat

```{r}
knitr::include_graphics("../figs/habitat.svg")
```

---

## Initial abundance distribution

```{r}
knitr::include_graphics("../figs/initial_abundance.svg")
```

---

## Preferential sampling

```{r}
knitr::include_graphics("../figs/commercial_effort.svg")
```

---

## Counter-preferential sampling

```{r}
knitr::include_graphics("../figs/station_prob.svg")
```

---

## Counter-preferential sampling - 50 survey stations

```{r}
knitr::include_graphics("../figs/survey_effort_50.svg")
```

---

## Counter-preferential sampling - 100 survey stations

```{r}
knitr::include_graphics("../figs/survey_effort_100.svg")
```

---

## Counter-preferential sampling - 200 survey stations

```{r}
knitr::include_graphics("../figs/survey_effort_200.svg")
```

---

## Counter-preferential sampling - 300 survey stations

```{r}
knitr::include_graphics("../figs/survey_effort_300.svg")
```

---

## Counter-preferential sampling - 400 survey stations

```{r}
knitr::include_graphics("../figs/survey_effort_400.svg")
```

---

## Abundance series

```{r}
knitr::include_graphics("../figs/abundance_series.svg")
```

---

## Estimation model

$$\log \vec{n} = \vec{X}\vec{\beta} + \vec{\Gamma}\vec{\lambda} + \vec{A}_{\omega}\vec{\omega}$$

- $n$: biomass
- $\vec{X}\vec{\beta}$: Year effects
- $\vec{\Gamma}\vec{\lambda}$: Vessel effects
- $\vec{A}_\omega\vec{\omega}$: Spatial random effects

--

$$c_{i} \sim \operatorname{Tweedie}(n_{i}, \phi, p)$$

- $c_{i} \ge 0$: Catch biomass observations
- $\phi > 0$: Dispersion parameter
- $1 < p < 2$: Shape parameter

---

## Estimation model

$$\hat{I}_{y} = \sum_{s \in \mathcal{S}} w_{s} \exp(\vec{X}^{*}_{y}\hat{\vec{\beta}} + \vec{A}^{*}_{\omega} \hat{\vec{\omega}})$$

- $\hat{I}_y$: Estimated index for year $y$
- $\mathcal{S}$: Spatial domain
- $\vec{X}^*\hat{\vec{\beta}}$: Projected year effect
- $\vec{A}^*_\omega\hat{\vec{\omega}}$: Projected spatial effect

---

## Simulation study

- 25 replicates
- Linear abundance commercial location preference
- 50, 100, 200, 300, or 400 survey stations
- Commercial fishing effort constant at 2000
- 500, 1000, or 2000 commercial observations used in estimation model

---

## Evaluation

- Index bias
- Index RMSE
- Index coverage
- Vessel effect estimates
- Vessel effect variance

---

# Results

---

## Index bias

```{r}
knitr::include_graphics("../figs/bias_plot2.svg")
```

---

## Index RMSE

```{r}
knitr::include_graphics("../figs/rmse_plot.svg")
```

---

## Index calibration

```{r}
knitr::include_graphics("../figs/calibration.svg")
```

---

## Vessel effect estimate

```{r}
knitr::include_graphics("../figs/lambda_val.svg")
```

---

## Variance of vessel effect

```{r}
knitr::include_graphics("../figs/lambda_sd.svg")
```

---

# Conclusions
 
- Bias can be reduced through counter-preferential sampling

- Vessel effects may be biased under preferential sampling


## Further work

- Optimal counter-preferential sampling designs

- Designs robust against shifting commercial effort distribution

- Estimate selectivity differences

- Multispecies settings
